Functional trait data tell new story about biodiversity change

### 2020-10-15 paper/manuscript.Rmd

Biodiversity timeseries were obtained from the Biotime database, a
global repository of assemblage timeseries with 386 individual studies
from multiple taxa.

-   matched all species to an ID using taxadb, if more than one species
    name matched to the same ID they were considered the same species
    and observations of those species were consolidated into one - only
    include time series with at least 4 species and at least 4 years

-   drop traits that were the same for all observed species

-   all non-categorical or binary traits were z-score scaled to give
    each trait equal weight
    ([@lepsQuantifyingInterpretingFunctional2006],
    [@schleuterUserGuideFunctional2010])

@@

@@

-   weight by abundance if available and biomass if not

@@

>>>>

Biodiversity timeseries were obtained from the Biotime database, a
global repository of assemblage timeseries with 386 individual studies
from multiple taxa. Timeseries were paired with their functional trait
data based on the predominant taxa in the timeseries. Due to limitations
in trait data and processing feasability, we included only mammal, bird,
and amphibian timeseries. Trait data were obtained from Elton Trait
Database (mammals and birds; [@wilman2014]) and Amphibio (amphibians,
[@oliveira2017]).

-   something here describing where the data in those databases comes
    from, that they're species-level means, etc**;**

-   also more description of how biotime is structured (studies, some
    with plots, etc)

In order to ensure taxonomic consistency across datasets, biotime
species were paired with trait data based on their species identifier
from the Integrated Taxonomic Information System database (citation),
obtained through the taxadb R package
([@normanTaxadbHighperformanceLocal2020]). If more than one species in
the assemblage data resolved to the same identifier, observations were
considered the same species. For trait data, traits for all species of
the same identifier were averaged. Only studies with at least 85% trait
coverage were included. In order to have a sufficient number of
observation years to observe a trend and a sufficient number of species
to calculate functional diversity metrics, studies with fewer than 4
years and years with fewer than 4 species observed were also excluded.

Functional diversity metrics were calculated for total yearly abundance.
For many studies there were a different number of samples within each
year. To adjust for this inconsistency in sampling effort we used
sample-based rarefaction by bootstrap resampling within year for each
timeseries based on the smallest number of samples in a year for that
timeseries.

@@

#### *Functional diversity metrics*

We calculated yearly metrics of functional diversity and turnover for
each plot within each study. All available trait data for each study was
included in the trait matrix with the exception of traits that were the
same value for all observed species in the study. All continuous traits
were z-score scaled to give each trait equal weight in the trait space
([@lepsQuantifyingInterpretingFunctional2006],
[@schleuterUserGuideFunctional2010]).

We used the function dbFD from the FD package (citation) to calculated
PCA-scaled trait matrices for each species in a timeseries and
functional diversity (FD), functional evenness (FEve), functional
diveregence (FDiv), and functional dispersion (FDis) based on those
trait matrices. The number of trait axes was limited to the maximum
number of traits that fulfills the criteria $s >= s^t$, where $s$ is the
number of species and \$t\$ is the number of traits. This limitation
allows for a sufficient number of axes to capture the trait space while
maintaining computational feasibility for metric calculations. Metrics
incorporated weighting based on either species abundance or biomass
based on the methods for each study. The PCA-scaled trait matrices were
then used as input to the BAT package....

@@

@@

#### *Analysis*
